---
title: "Section 2: ggplot2"
author: "Matt Higham"
format: 
  html:
    embed-resources: true
---

```{r}
library(tidyverse)
stat113_df <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/stat113.csv")
```

## Section 2.3: Introduction to Plotting

One quantitative variable.

```{r}
stat113_df
ggplot(data = stat113_df, aes(x = GPA)) +
  geom_histogram(colour = "navyblue", fill = "plum2", bins = 15)

ggplot(data = stat113_df, aes(x = GPA)) +
  geom_freqpoly(bins = 17)

ggplot(data = stat113_df, aes(x = GPA)) +
    geom_histogram(colour = "black", fill = "white", bins = 17) +
    geom_freqpoly(bins = 17)
```

One categorical variable.

```{r}
ggplot(data = stat113_df, aes(x = SocialMedia)) +
  geom_bar(colour = "mistyrose", fill = "lightcoral")
```

Two categorical variables.

```{r}
ggplot(data = stat113_df, aes(x = Sport, fill = SocialMedia)) +
  geom_bar() 

ggplot(data = stat113_df, aes(x = Sport, fill = Award)) +
  geom_bar(position = "fill") +
  labs(y = "Proportion")
```

One categorical and one quantitative

```{r}
ggplot(data = stat113_df, aes(x = Award, y = Exercise)) +
  geom_boxplot()

ggplot(data = stat113_df, aes(x = Sex, y = GPA)) +
  geom_violin(colour = "slategrey", fill = "yellow")
## there no evidence of association between sex and gpa

ggplot(data = stat113_df, aes(x = Award, y = Exercise)) +
  geom_jitter(width = 0.10)
## the strip plot shows a rough idea of sample size
```

```{r}
library(tidyverse)
fitness_full <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/higham_fitness_clean.csv",
                         col_types = list(stepgoal = col_factor())) |>
  mutate(weekend_ind = case_when(weekday == "Sat" | weekday == "Sun" ~ "weekend",
                                 .default = "weekday"))
```

## Section 2.4: Practice

```{r}
stat113_restructured <- stat113_df |> group_by(SocialMedia) |>
  summarise(n_social = n())
stat113_restructured

ggplot(data = stat113_restructured, aes(x = SocialMedia)) +
  geom_bar()
## each level of socialmedia appears only once so each
## bar gets a height of one

ggplot(data = stat113_restructured, aes(x = SocialMedia,
                                        y = n_social)) +
  geom_col()
```

Class Exercise 2. 

```{r}
fitness_full
ggplot(data = fitness_full, aes(x = distance, y = active_cals)) +
  geom_point() +
  geom_smooth()
```

geom_point() and also geom_smooth()

Class Exercise 3. geom_line() and also geom_smooth()

```{r}
ggplot(data = fitness_full, aes(x = Date, y = steps)) +
  geom_line() +
  geom_smooth()
## the line plot is appropriate here because there is only one
## y value (steps) for each unique x value (Date). This is often
## true when x is time.
```

Class Exercise 4. use either colour or faceting or both.

```{r}
ggplot(data = stat113_df, aes(x = Exercise, y = GPA)) +
  geom_point() +
  facet_wrap(~ Sex)
ggplot(data = stat113_df, aes(x = Exercise, y = GPA)) +
  geom_point() +
  facet_wrap(~ Sport)

ggplot(data = stat113_df, aes(x = Exercise, y = GPA, colour = Sex)) +
  geom_point() +
  facet_wrap(~ Sport)

ggplot(data = stat113_df, aes(x = Exercise, y = GPA)) +
  geom_point() +
  facet_grid(Sex ~ Sport)
## facet_grid() is generally useful for faceting by two categorical
## variables.
```

Class Exercise 5.

```{r}
ggplot(data = stat113_df, aes(x = Award, y = Exercise)) +
  geom_boxplot(outlier.shape = 8)
```

## Your Turn

Your Turn 1. 

a. 25 rows (one row for each student that we are getting information on) and only one column would be needed (Major).

| Major |

| Statistics |
|Statistics |
| Other |
| Other |
.....
| Data Science |

b. 3 rows (one for each levele of Major) and two columns (one for Major and one that has the counts).

|Major | count |
| Data Science  | 8 |
| Other | 10 |
|Statistics | 7 |

Your Turn 2.

```{r}
ggplot(data = stat113_df, aes(x = GPA, y = Wgt,
                              colour = SocialMedia)) +
  geom_point()
ggplot(data = stat113_df, aes(x = GPA, y = Wgt)) +
  geom_point() +
  facet_wrap(~ SocialMedia)

## in general, if the categorical variable has
## few levels (2-4), colour often works better.

## if there are a lot of levels (like more than 7), colours become hard to distinguish and it's better to facet.
```

