---
title: "Section 3: dplyr"
author: "Matt Higham"
format: 
  html:
    embed-resources: true
---

## Section 3.1

```{r}
library(babynames)
library(tidyverse)
babynames
```

## Section 3.2: Practice

Class Exercises 1 and 2.

```{r}
babynames_m10 <- babynames |> filter(sex == "M" & year == 2017) |>
  arrange(desc(n)) |>
  slice(1:10)

ggplot(data = babynames_m10, aes(x = name, y = n)) +
  geom_col()
```

Class Exercises 3 and 4.

```{r}
babynames_matthew <- babynames |> filter(name == "Matthew" & sex == "M")

ggplot(data = babynames_matthew, aes(x = year, y = n)) +
  geom_line() +
  geom_vline(xintercept = 1991, linetype = 2) ## extra
```

Class Exercise 5. Piping only works if the first argument of the function you are piping into is a data frame. 

Your Turn 1.

```{r}
babynames |> filter(year != 1945)

babynames |> filter(n >= 400 & n <= 5000)
```

Your Turn 2.

```{r}
babynames_mh <- babynames |> filter((name == "Monica" & sex == "F") | (name == "Hilary" & sex == "F")) 

ggplot(data = babynames_mh, aes(x = year, y = n, colour = name)) +
  geom_line()
```

Your Turns and 3 and 4.

```{r}
babynames_5 <- babynames |> filter((name == "Matthew" & sex == "M") |
                      (name == "Jessica" & sex == "F") |
                      (name == "Ivan" & sex == "M") |
                      (name == "Robin" & sex == "M") |
                      (name == "Lillian" & sex == "F"))
ggplot(data = babynames_5, aes(x = year, y = n, colour = name)) +
  geom_line()
```

## Section 3.4

Class Exercise 1.

```{r}
library(tidyverse)
slumajors_df <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/SLU_Majors_17_21.csv")
slumajors_df

slumajors_df <- slumajors_df |>
  mutate(ntotal = nfemales + nmales) |>
  mutate(percfemale = 100 * nfemales / (nfemales + nmales))

slumajors_df <- slumajors_df |>
  mutate(morewomen = if_else(percfemale > 50,
                             true = "Yes",
                             false = "No")) |>
  mutate(large_majority =
           case_when(percfemale >= 70 ~ "female",
                     percfemale <= 30 ~ "male",
                     percfemale > 30 & percfemale < 70 ~ "none")) 


slumajors_df |> mutate(testcase = case_when(percfemale >= 40 ~ "Yes",
                                            percfemale > 60 ~ "No")) |>
  relocate(testcase) ## moves testcase so that it appears first
## when a row satisfies multiple conditions, case_when() will go
## with whatever condition is firat
## when a row satisfies no conditions, the new variable is NA
```

Class Exercise 2.

```{r}
library(babynames)

babynames |> 
  group_by(year, sex) |>
  mutate(rank_pop = rank(desc(n))) |>
  filter(rank_pop == 1) |>
  ungroup() |>
  group_by(name) |>
  summarise(n_pop = n()) |>
  arrange(desc(n_pop))
```

Class Exercise 3.

```{r}
babynames |> 
  group_by(year, sex) |> 
  arrange(desc(n)) |>
  slice(1) |>
  ungroup() |>
  group_by(name) |>
  summarise(n_pop = n()) |>
  arrange(desc(n_pop))

babynames |> 
  group_by(year, sex) |> 
  filter(n == max(n)) |>
  ungroup() |>
  group_by(name) |>
  summarise(n_pop = n()) |>
  arrange(desc(n_pop))
```

Class Exercise 4.

```{r}
babynames_test <- babynames |>
  group_by(year, sex) |> mutate(ntest = n / prop)
babynames_test |> slice(1, 2, 3, 4, 5)
## gets first 5 rows for each group

babynames_test |> ungroup() |> slice(1, 2, 3, 4, 5)
```

Class Exercise 5.

```{r}
fitness_df <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/higham_fitness_clean.csv") |>
  select(-stepgoal)
fitness_df |>
  mutate(stepgoal_ind = if_else(steps >= 10000,
                                true = 1,
                                false = 0)) |>
  relocate(stepgoal_ind) |>
  group_by(year) |>
  summarise(prop_goal = mean(stepgoal_ind)) |>
  arrange(desc(prop_goal))
```

## Your Turn

Your Turn 1. Skip.

Your Turn 2.

```{r}
toy_df <- tibble(x = c(NA, 3, 4, 7),
                 y = c(1, 4, 3, 2),
                 z = c("A", "A", "B", NA))

## a
toy_df |> mutate(new_x = x + 5)
## the NA value is still NA

## b
toy_df |> arrange(desc(x))
## NAs go at the bottom

## c
toy_df |> filter(x < 5)
## NAs get dropped

## d
toy_df |> summarise(x_mean = mean(x))
## return an NA (the mean of 3, 4, 7, and NA is also NA)

toy_df |> summarise(x_mean = mean(x, na.rm = TRUE))
## can remove missing values first

toy_df |> group_by(z) |> summarise(x_mean = mean(x))
## treats NAs as their own group.

```

Your Turn 3.

a. mutate() is safe: we are just adding a new variable without modifying any of the original data.

b. arrange() is also safe: the data is still the same: just in a different order.

c. filter() is less safe: you are removing rows that are then gone for good.

d. summarise() is not safe at all: it removes columns that aren't involved, it reduces the number of rows to one row per group.

e. select() is less safe: would only do this for columns that you _know_ you will never need.

Your Turn 4.

```{r}
slumajors_df |>
  summarise(meantotalmajor = mean(ntotal),
            totalgrad = sum(ntotal)) 
slumajors_df |>
  mutate(meantotalmajor = mean(ntotal),
            totalgrad = sum(ntotal)) |>
  select(meantotalmajor, totalgrad, everything())
## mutate() adds a new column, keeps the same number of rows and keeps
## all of the other data.

## summarise() makes a new table that is one row per group (so just 1 row here)
```

Your Turn 5.

```{r}
library(babynames)
babynames_distinct <- babynames |> group_by(year) |>
  summarise(n_distinct = n())

ggplot(data = babynames_distinct, aes(x = year, y = n_distinct)) +
  geom_line()
```

Your Turn 6.

```{r}
babynames_fm <- babynames |> group_by(year, sex) |>
  summarise(n_distinct = n())
babynames_fm

ggplot(data = babynames_fm, aes(x = year, y = n_distinct)) +
  geom_line(aes(colour = sex))
```

Your Turn 7.

```{r}
babynames |> group_by(name) |>
  summarise(total_births = sum(n)) |>
  arrange(desc(total_births))
  
```

