---
title: "Section 4: Communication"
author: "Matt Higham"
execute:
  warning: false
format: 
  html:
    embed-resources: true
---

Class Exercises

```{r}
#| echo: false
library(tidyverse)
```

```{r}
library(gapminder)
country_df <- gapminder
glimpse(country_df)

country_df
ggplot(data = country_df, aes(x = year, y = lifeExp, group = country)) +
  geom_line() +
  facet_wrap(~ continent)

## step 1: figure out what these three countries are.

country_df |> group_by(continent) |>
  arrange(lifeExp) |>
  slice(1)
## alt: group_by(continent) |> filter(lifeExp == min(lifeExp))
## Africa: Rwanda
## Europe: Turkey
country_df |> filter(continent == "Asia" & year <= 1980 & year >= 1970 & lifeExp < 35)
## Asia: Cambodia

## colour the three countries with the "interesting" patterns
## hint: use a second geom_line()
country_3 <- country_df |> filter(country == "Rwanda" | country == "Cambodia" | country == "Turkey")

## change x and y axis labels
## use a different theme
## use a CVD-friendly colour scale
## add a title and a caption, where the caption 
## gives the events in Rwanda and Cambodia
## look into alpha, a parameter that can be used in geoms to 
## control transparency
ggplot(data = country_df, aes(x = year, y = lifeExp, group = country)) +
  geom_line(alpha = 0.2) +
  geom_line(data = country_3, aes(colour = country)) +
  facet_wrap(~ continent) +
  labs(x = "year", y = "life expectancy",
       title = "Life Expectancy through Time",
       caption = "Cambodia and Rwanda had a genocide and a civil war.") +
  scale_colour_viridis_d() +
  theme_minimal()

## alternative: label the three country names directly 
## at their minimum life expectancies

country_3_lab <- country_df |> filter(country == "Rwanda" | country == "Cambodia" | country == "Turkey") |> 
  group_by(country) |>
  filter(lifeExp == min(lifeExp))
  
ggplot(data = country_df, aes(x = year, y = lifeExp, group = country)) +
  geom_line() +
  geom_text(data = country_3_lab, aes(label = country)) +
  facet_wrap(~ continent)
```

Section 4.4

Your Turn 1. Changing the y-axis limits and labels with point-and-click makes it hard for someone else to reproduce the exact same graph. Removing the outliers is also not reproducible unless he leaves a very detailed written description. Finally, they use a t-test function, which they would need to describe in detail to make this reproducible.

Your Turn 2.

```{r}
library(tidyverse)
fitness_full <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/higham_fitness_clean.csv",
                         col_types = list(stepgoal = col_factor())) |>
  mutate(weekend_ind = case_when(weekday == "Sat" | weekday == "Sun" ~ "weekend",
                                 .default = "weekday"))

## 2.4 class exercise 3.
fitness_lab <- fitness_full |>
  filter(Date == "2019-04-14" | Date == "2021-05-24") |>
  mutate(event = c("half-marathon", "grand canyon"))
## creating a new variable with values that we supply
fitness_lab |> relocate(event)

ggplot(data = fitness_full, aes(x = Date, y = steps)) +
  geom_line() +
  geom_smooth() +
  geom_text(data = fitness_lab, aes(label = event))
```

Your Turn 3.

