---
title: "Section 7: Factors"
author: "Matt Higham"
format: 
  html:
    embed-resources: true
---

```{r}
library(tidyverse)
library(here)
pokemon_df <- read_csv(here("data/pokemon_allgen.csv")) |>
  mutate(Generation_cat = factor(Generation))
```

Class Exercise 1. This is not appropriate since Defense is a quantitative variable (so we cannot see things like spread/variability, shape, outliers, sample size).

These plots are appropriate for things like counts and totals of a categorical variable.

Class Exercise 2. group_by(#) |> slice(1) |> ungroup() This works because we set up groups for each unique # and then slicing (1) keeps the first row for each unique #. 

Alternatively:

```{r}
pokemon_df |> distinct(`#`, .keep_all = TRUE)
## keeps one row per unique value(s) of the variable(s) you provide
## equivalent to the slice(1) approach above.
```

Class Exercise 3.

```{r}
pokemon_nodup <- pokemon_df |> group_by(`#`) |> slice(1) |>
   ungroup()
pokemon_long <- pokemon_nodup |>
  pivot_longer(c(`Type 1`, `Type 2`),
               names_to = "Number",
               values_to = "Type")

pokemon_long <- pokemon_long |>
  filter(!is.na(Type)) |>
  mutate(Type_Deford = fct_reorder(.f = Type, .x = Defense,
                                   .fun = median))
ggplot(data = pokemon_long, aes(x = Type_Deford,
                               y = Defense)) +
  geom_boxplot() + 
  coord_flip()
## in general, forcats functions (especially fct_reorder() 
## do not really work well with grouped data frames - so
## will need to ungroup before running fct_reorder ).
```

Your Turn

```{r}
library(tidyverse)
library(here)
standings_df <- read_csv(here("data/standings.csv"))
standings_df |> relocate(playoffs)
```

Your Turn 1.

```{r}
table(standings_df$team_name)
```

Your Turns 2 through 4
```{r}
standings_df <- standings_df |> mutate(team_name = fct_recode(team_name,
                                              Commanders = "Redskins")) |>
  unite("franchise", c(team, team_name), sep = " ")

standings_clean <- standings_df |>
  mutate(franchise = fct_collapse(franchise, `St. Louis LA Rams` = c("St. Louis Rams", "Los Angeles Rams"),
                                  `San Diego LA Chargers` = c("San Diego Chargers", "Los Angeles Chargers")))
standings_clean |> relocate(playoffs)

```

Your Turn 5.

```{r}
## one strategy: start with filtering to keep only the rows where the team made the playoffs
## then we could group_by franchise and summarise to count up the number of times each
## team appears in the playoffs data set.
```

