---
title: "Section 10: Joining"
author: "Matt Higham"
format: 
  html:
    embed-resources: true
---

Class Exercise 0.

a. 5 rows

b. 6 rows.

c. 8 rows.

d. 3 rows (A, C, E).

e. 3 rows.

f. 2 rows.


Class Exercise 1. To use bind_cols(), you must be sure that the order of the cases in each of the two data sets matches exactly.

Class Exercise 2.

```{r}
library(tidyverse)
library(here)
slumajors_df <- read_csv(here("data/SLU_Majors_17_21.csv"))
collegemajors_df <- read_csv(here("data/college-majors.csv"))
glimpse(slumajors_df)
glimpse(collegemajors_df)

left_join(slumajors_df, collegemajors_df, join_by(Major == Major))
## this does not work since the values for the keys in the two data
## sets do not have the same syntax (title case vs. all caps).
```


Class Exercise 3.

```{r}
library(tidyverse)
library(here)
mortality_df <- read_csv(here("data/gun_violence_us.csv"))
mortality_df

dc_df <- tibble(state = "DC", mortality_rate = 16.7,
                ownership_rate = 0.087, region = "NE")
dc_df

all_df <- bind_rows(mortality_df, dc_df)
```

Class Exercises 4 through 8.

```{r}
df1 <- as_tibble(state.x77)
df2 <- as_tibble(state.abb)
df1
df2

## assuming that df1 and df2 have the same state order
states_df <- bind_cols(df1, df2)

## ce 5
left_join(all_df, states_df, join_by(state == value))
## OR
full_join(all_df, states_df, join_by(state == value))

## ce 6
left_join(states_df, all_df, join_by(value == state))
## OR
inner_join(states_df, all_df, join_by(value == state))

## ce 7
ne_df <- all_df |> filter(region == "NE")
semi_join(states_df, ne_df, join_by(value == state))

## ce 8
notne_df <- all_df |> filter(region != "NE")
anti_join(states_df, notne_df, join_by(value == state))
```



Your Turn 1.


a. true

b. true. the full_ and left_ join will return at least the same rows as an inner_join and possibly others. 

c. false. see Your Turns a and e for a counterexample.

A, 5
D, 6


A, 1
A, 0
A, 4
B, 5
B, 6
C. 6

inner_join:
A, 5, 1
A, 5, 0
A, 5, 4

semi_join:
A, 5





Your Turn 2.

a. 6 rows.

b. 8 rows (A, A, B, C, C, D, D, D)

c. 10 rows (A, A, B, C, C, D, D, D, F, F)

d. 4 rows (A, A, C, C)

e. 4 rows (A, A, C, C, yvar also gets dropped).

f. 2 rows (F, F).

Your Turn 4.

left_join(courses_df, info_df): 300 rows.

semi_join(info_df, courses_df): 5 rows (keep the rows corresponding to the 5 profs that are in courses_df).

anti_join(courses_df, info_df). 285 rows (300 rows minus 15 rows that have the matching key in info_df).

inner_join(courses_df, info_df). 15 rows (3 times 5: 15 rows for the 5 profs that appear in both data set- each prof has 3 courses).

full_join(courses_df, info_df): 310 rows (15 from the inner_join plus the 285 from the coureses df without the matching key in info_df plus the 10 rows in info_df without the matching key in courses_df).
