---
title: "Section 12: Intro to Stat. Learning with knn"
author: "Matt Higham"
format: 
  html:
    embed-resources: true
---

```{r}
set.seed(1119)
library(tidyverse)
library(pander)
library(here) 

pokemon <- read_csv(here("data", "pokemon_full.csv")) |>
  ## keeping the types we are interested in
  filter(Type %in% c("Steel", "Dark", "Fire", "Ice")) |>
  mutate(across(where(is.numeric), ~ (.x - min(.x)) /
                                 (max(.x) - min(.x)))) |>
  ## scaling our candidate predictors so that they are all between 0 and 1
  ## where(is.numeric) says to only apply the formula to columns that are numeric.
  mutate(Type = as.factor(Type))
pokemon
```

```{r}
train_sample <- pokemon |>
  slice_sample(n = 75)
## randomly select 75 to be in our training sample
test_sample <- anti_join(pokemon, train_sample)
## all other pokemon (45) are in the test sample.
```

```{r}
library(GGally)
ggpairs(data = train_sample, columns = c(7, 8, 12, 3))
## Defense looks like it will be useful to include in
## an initial knn model because Steel types have higher
## defense, on average, than the other types (not a lot of overlap in the boxplots)
## HP and Attack might be useful, might not: we will see
## once we know how to formally evaluate model performance.
```

```{r}
library(kknn)

## fit the knn model with 9 nearest neighbors
knn_mod <- kknn(Type ~ Attack + Defense + Speed + SpAtk, train = train_sample,
                test = test_sample, k = 11)
knn_mod

knn_mod$fitted.values
## these are the classifications for each pokemon in the test_sample
## according to the model. saying the first is classified as Fire, second
## is Fire, ...., fifth is Ice, etc.
```

```{r}
## taking our model predictions first and then the actual pokemon types second
table(knn_mod$fitted.values, test_sample$Type) 
```

Exercise 2. There are 12 pokemon in our test sample that the model classified as Fire and were actually Fire.

Exercise 3. There are 3 pokemon that the model predicted to be ICe but are actually Dark.

Exercise 4. There are 0 pokemon that the model predicted to be Steel that were actually Dark.

Exercise 5. classification rate for this model is: (2 + 12 + 1 + 5) / 45

```{r}
(2 + 12 + 1 + 5) / 45
## classification rate for this model is 0.444

## want to make this as high as we possibly can.

## code to automate this:
tab <- table(knn_mod$fitted.values, test_sample$Type) 
sum(diag(tab)) / sum(tab)
## SpAtk increased the classification rate to 0.47 so it seems to be useful.
```

Exercise 7.

```{r}
## with the training data, figure out the most common Type
train_sample |> group_by(Type) |>
  summarise(n_type = n()) |>
  arrange(desc(n_type))
## Fire is the most common type in the training sample

## with the test data, figure out the proportion of Pokemon that 
## have that most common Type in the training. This is our baseline
## classification rate.
test_sample |> mutate(type_ind = if_else(Type == "Fire",
                                         true = 1, false = 0)) |>
  relocate(type_ind) |>
  summarise(prop_fire_test = mean(type_ind))
## baseline classification rate: 0.356
## our model at least did better than the baseline!
```


Exercise 8. Making k too small might mean that you leave your classifications up to chance (high variance model). but making k too large might mean that you are using pokemon that are not that related to your test pokemon (high bias model). 

## Section 12.4

```{r}
library(tidyverse)

pokemon <- read_csv(here::here("data/pokemon_full.csv")) |>
  mutate(Type = as.factor(Type))
set.seed(1119)

## scale the quantitative predictors
pokemon_scaled <- pokemon |>
  mutate(across(where(is.numeric), ~ (.x - min(.x)) /
                  (max(.x) - min(.x))))

train_sample <- pokemon_scaled |>
  slice_sample(n = 550)
test_sample <- anti_join(pokemon_scaled, train_sample)

library(kknn)

knn_mod <- kknn(Type ~ HP + Attack + Defense + Speed + SpAtk + SpDef + height + weight,
                train = train_sample, test = test_sample, k = 9)
knn_mod

tab <- table(knn_mod$fitted.values, test_sample$Type)
sum(diag(tab)) / sum(tab)
```

```{r}
get_class_rate <- function(k_val) {
  knn_mod <- kknn(Type ~ HP + Attack + Defense + Speed + SpAtk + SpDef + height + weight,
                  train = train_sample, test = test_sample, k = k_val)
  
  tab <- table(knn_mod$fitted.values, test_sample$Type)
  class_rate <- sum(diag(tab)) / sum(tab)
  
  return(class_rate)
}
## one input called k_val
## returns a classification rate
get_class_rate(k_val = 20)

k_vec <- 1:120
class_rates <- map_dbl(k_vec, get_class_rate)
## run get_class_rate with k_val = 1, then
## run get_class_rate with k_val = 2, then ....
## run get+class_rate with K_val = 120

class_rate_df <- tibble(k_vec = k_vec, class_rates = class_rates)
class_rate_df

## construct a line plot looking at class_rate for different ks
## write dplyr code to give you the k_value with the max classification rate.
ggplot(data = class_rate_df, aes(x = k_vec, y = class_rates)) +
  geom_line() +
  theme_minimal()
class_rate_df |> filter(class_rates == max(class_rates))
## may choose k = 64
```

