---
title: "Section 16: Introduction to SQL"
author: "Matt Higham"
format: 
  html:
    embed-resources: true
---

```{r}
library(RSQLite)
library(DBI) 
library(dplyr)
library(dbplyr)

imdb_db <- dbConnect(RSQLite::SQLite(), 
                     dbname = here::here("data/imdb_tv.sqlite"))

dbListTables(conn = imdb_db)
## show all of the data tables (data frames) stored in this database

dbListFields(conn = imdb_db, name = "episode_rating")
## shows the fields (variables) in the episode_rating table

library(tidyverse)

sql <- "
  SELECT parentTconst, averageRating, numVotes
  FROM show_ratings 
  WHERE numVotes > 100000
"

dbGetQuery(imdb_db, sql) |>
  as_tibble()

## exercises 1 and 2
## WHERE is similar to dplyr's filter()
## SELECT is similar to dplyr's select()
## FROM just tells SQL what data table we want to perform these
## operations on.

```

## Section 16.2: `dbplyr`

```{r}
library(dbplyr)
episode_rating_tbl <- tbl(imdb_db, "episode_rating")
episode_rating_tbl
## store the episode_rating data table as its own object
## so that we can work with it in dbplyr

## in this printout, we see the database location
## we also see ?? for the number of rows because SQL
## only runs the query on a subset of the data.
```

```{r}
episode_query1 <- episode_rating_tbl |> 
  filter(averageRating > 9.0) |> 
  select(parentTconst, Episode_Name, numVotes, averageRating)
episode_query1

episode_query1 |> collect()
## converts the data table to a tibble in R
## and queries all of the rows.

episode_query1 |>
  show_query()

mean_query1 <- episode_rating_tbl |>
  group_by(parentTconst) |>
  summarise(mean_rating = mean(averageRating)) |>
  arrange(desc(mean_rating))
mean_query1

mean_query1 |> show_query()
```

Exercise 3.

```{r}
n_show_query <- episode_rating_tbl |>
  group_by(parentTconst) |>
  summarise(n_shows = n())
n_show_query |> show_query()

## group_by() in dplyr is similar to GROUP BY in SQL
## summarise() in dplyr is similar to SELECT, where
## COUNT(*) is like n() and then AS assigns a name to the summary variable
```

## Section 16.3: SQL Connections to `dplyr`

* `SELECT` in SQL matches with:
    * `select()`
    * `rename()` (in SQL, SELECT OLDNAME AS NEWNAME)
    * `mutate()` (in SQL, SELECT x + y AS xplusy)
    * `summarise()` (IN SQL SELECT MEAN(x) AS mean_of_x)
* `GROUP BY` in SQL matches with `group_by()`.
* `WHERE` in SQL matches with `filter()`.
* `ORDER BY` in SQL matches with `arrange()`.

```{r}
show_info_tbl <- tbl(imdb_db, "show_info")

left_join(episode_rating_tbl, show_info_tbl,
          join_by(parentTconst == parentTconst)) |>
  show_query()
```

In SQL, the key is specified as ON (datatable1.key = datatable2.key) instead of `left_join(datatable1, datatable2, join_by(key == key))`

Exercise 4.

```{r}
show_info_tbl |>
  filter(Show_Name == "Ozark") |>
  show_query()
```

Class Exercise 1.

```{r}
show_info_tbl |>
 ## select(startYear, parentTconst, Show_Name) |>
  rename(show_name = Show_Name) |>
  arrange(desc(startYear)) |>
  show_query()
```

Your Turns 1 through 4.

```{r}
episode_rating_tbl |>
  group_by(parentTconst) |>
  summarise(best_episode = max(averageRating, na.rm = TRUE)) |>
  show_query()
## summary functions in SELECT (MAX, MIN, MEDIAN, MEAN, COUNT, etc.)
## get translated to summarise() in  dplyr.
## things in SELECT like x * y or (z - w) / w get translated
## to mutate() in dplyr.
```


Your Turn 2.


```{r}
#| error: true
episode_rating_tbl |> mutate(test = fct_reorder(parentTconst,
                                                averageRating))
## cannot use lubridate, forcats, stringr, or other r specific
## packages on database tables unless you collect() it first.
```


Your Turn 3.

```{r}
episode_rating_tbl |> filter(seasonNumber != "1") |>
  show_query()
## != is the same in both

episode_rating_tbl |>
  filter(seasonNumber %in% c("1", "2")) |>
  show_query()
## %in% gets translated to IN in SQL

episode_rating_tbl |> distinct(parentTconst) |>
  show_query()
## distinct() gets translated to DISTINCT in SQL
```

YouR Turn 4.

```{r}
episode_rating_tbl |>
  filter(seasonNumber == "1") |>
  group_by(parentTconst) |>
  summarise(n_episodes = n()) |>
  arrange(desc(n_episodes)) |>
  show_query()
```

```{r}
imdb_df <- left_join(episode_rating_tbl, show_info_tbl,
          join_by(parentTconst == parentTconst)) |>
  collect()
imdb_df$Show_Name |> unique()

imdb_stranger <- imdb_df |> filter(Show_Name == "Stranger Things") |>
  mutate(episodeNumber = as.numeric(episodeNumber))

ggplot(data = imdb_stranger, aes(x = episodeNumber, y = averageRating)) +
  geom_line() +
  facet_wrap(~ seasonNumber)
```


